\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{hyperref}

\title{Noetheris: A Local-First Structural Kernel for Quantum-Inspired Security Analysis}
\author{Mayckon Giovani}
\date{v0.2 working paper}

\begin{document}

\maketitle

\begin{abstract}
Noetheris is a local-first research kernel for representing security-critical
system structure as executable artifacts: structural intermediate
representations, QUBO/Ising models, Boolean oracle mappings, certificate
objects, and small formal-methods kernels. This paper defines the intended
v0.2 scope: a narrow external solver boundary for Ocean-compatible QUBO
exchange, Qiskit-compatible oracle export, and deterministic local replay of
solver candidates. The release does not claim quantum advantage, hardware
performance, cryptographic compromise, or production security certification.
\end{abstract}

\section{Problem}

Security failures in distributed and cryptographic infrastructure often arise
from structural conditions: unsafe state transitions, dependency cycles,
threshold-policy violations, migration blockers, and inconsistent evidence
about optimization results. Noetheris models those structures explicitly and
compiles bounded instances into representations that can be checked locally.

The v0.2 objective is deliberately small. The project should strengthen the
boundary between a deterministic local kernel and optional external solver
ecosystems. The central rule is that external solver output is not trusted as a
result until Noetheris replays the witness locally.

\section{Structural Model}

Let a structural instance be
\[
  \mathcal{S} = (V, E, C, I, O, A, M),
\]
where $V$ is a typed node set, $E$ is a typed edge set, $C$ is a set of
constraints, $I$ is a set of invariants, $O$ is a set of objectives, $A$ is a
set of adversary actions, and $M$ is metadata. Canonical JSON serialization
defines a deterministic problem hash
\[
  h_P = H(\mathrm{canonical}(\mathcal{S})).
\]
This hash is the stable identity of the source problem for replay and
certificate checks.

\section{QUBO And Ising Lowering}

Noetheris compiles bounded structural choices into binary variables
$x_i \in \{0,1\}$ and an energy function
\[
  E_Q(x) = c + \sum_i a_i x_i + \sum_{i<j} b_{ij} x_i x_j.
\]
The compiler canonicalizes duplicate, reversed, and self-quadratic terms before
export. A compiled model hash
\[
  h_C = H(\mathrm{canonical}(h_P, E_Q, \mathrm{metadata}))
\]
binds the source problem to the emitted optimization model.

The Ising lowering uses the convention $x_i = (1 - z_i)/2$ with
$z_i \in \{-1,1\}$. Energy equivalence is tested over small instances by
enumerating assignments.

\section{External Solver Boundary}

For Ocean and D-Wave-oriented workflows, Noetheris emits a structured QUBO
exchange artifact:
\[
  (X, c, L, Q, h_C),
\]
where $X$ is the ordered variable list, $c$ is the offset, $L$ is the linear
term list, $Q$ is the quadratic term list, and $h_C$ is the compiled model
hash. If Ocean is installed, the exchange can be materialized locally as a
\texttt{dimod.BinaryQuadraticModel}. This does not require a sampler, token, or
cloud account.

A solver candidate is a tuple
\[
  w = (h_P, h_C, x, \hat{E}, \mu),
\]
where $x$ is an assignment, $\hat{E}$ is the reported energy, and $\mu$ is
solver metadata. Noetheris accepts $w$ only if the hashes match, the assignment
domain is exact, and
\[
  E_Q(x) = \hat{E}
\]
within an explicit tolerance. Embedding metadata may be recorded when supplied
externally, but local Noetheris code does not infer hardware embedding quality.

\section{Circuit Oracle Boundary}

For gate-model workflows, a Boolean predicate $\phi$ over structural variables
is represented by the reversible oracle
\[
  O_\phi |x\rangle |y\rangle = |x\rangle |y \oplus \phi(x)\rangle.
\]
The v0.2 boundary focuses on small predicates with exact truth tables,
symbolic reversible gates, cleanup accounting, QASM-like text, and optional
local Qiskit circuit summaries. IBM Runtime execution is outside the default
path.

\section{Certificates}

Certificates bind the problem hash, compiled model hash, witness assignment,
energy terms, constraints, and replay metadata. The validation policy is
fail-closed: arithmetic mismatches, hash mismatches, unknown variables, missing
variables, and status inconsistencies reject the certificate or candidate.

\section{Reproducibility}

The expected local validation path is:
\[
\begin{array}{ll}
\text{audit} & \texttt{bash scripts/run\_audit.sh} \\
\text{examples} & \texttt{bash scripts/run\_examples.sh} \\
\text{benchmarks} & \texttt{bash scripts/run\_benchmarks.sh}
\end{array}
\]
Optional packages may enrich reports but must not change the trust boundary.

\section{Limitations}

Noetheris v0.2 remains a research engineering kernel. It does not benchmark
IBM Quantum or D-Wave hardware, does not claim quantum advantage, does not
assess deployed cryptography as compromised, and does not provide production
security certification. The implemented value is narrower: deterministic local
compilation, explicit solver boundaries, replayable witnesses, and honest
evidence artifacts.

\section{Conclusion}

The v0.2 release should improve Noetheris by making optional quantum and
annealing integrations more inspectable without weakening the security model.
The engineering invariant is simple: export may cross a solver boundary, but
verification remains local.

\end{document}
